---
title: "Thalamus divisions CCEP comparison"
output:
  html_document:
    df_print: paged
---

Note: Based on the antTH vs pstTH CCEP ITPC comparison, we can see distinct trends of ups and downs in the electrode wise comparisons, hence we hypothesize that there is a systematical clustering of the antTH-pstTH ITPC differences due to a functionally division within the anatomical regions of interest. Here we are testing this hypothesis for SM and PMC.

```{r setup, include=FALSE}
require(knitr)
require(plyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(hrbrthemes)
library(viridisLite)
require(reshape2)
library(tidyr)
library(gmodels)
library(fmsb)
library(RColorBrewer)
library(ggpubr)
library(vtable)
library(grid)
library(kableExtra)
library(factoextra)
library(pracma)
library(lme4)

opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
output_folder = '~/Dropbox/Stanford_Matters/data/THAL/CCEP/results/explore3_new2Sbjs'
figpath = '~/Dropbox/Stanford_Matters/data/THAL/CCEP/Plots/explore3'
opts_knit$set(root.dir = output_folder)

```


```{r}
setwd('~/Dropbox/Stanford_Matters/data/THAL/CCEP/results/explore3_new2Sbjs')
d = read.csv('table_CCEPnewpipOutput_wholebrain_anatomical_info_activationRedone.csv')

# further filter data
vnames = colnames(d)
peakMat = d[,grep('pks_time', vnames)]
#write.csv(d, na = "NaN", quote = FALSE, file = 'table_CCEPnewpipOutput_wholebrain_anatomical_info2_JPlabelsorted.csv')
#=================================================================================================
d0 <- d %>% filter((!JP_label_in %in% c("","empty","NAN","EXCLUDE", 'NA', 'INSULA/LFC')) &
  (!JP_label_out  %in% c("","empty","NAN","EXCLUDE", 'NA', 'INSULA/LFC')) & 
    !is.na(JP_label_in) & rCrossBorder == 0 & sCrossBorder == 0 & eudDist>5 &
    rowSums(peakMat, na.rm=TRUE) > 10
  ) 


#d <- d0 %>% filter(rowSums(peakMat, na.rm=TRUE) > 10) 
#head(d)

dinUnique <- d0 %>% mutate(elec_ID = paste(subject, record_chan)) %>% select(subject, aSubID, elec_ID, JP_label_in) %>% filter(!duplicated(elec_ID))

doutUnique <- d0 %>% mutate(elec_ID = paste(subject, stim_chan)) %>% select(subject, aSubID, elec_ID, JP_label_out) %>% filter(!duplicated(elec_ID))

```

### check electrode coverage for the recorded channels across JP_label and subjects
```{r}
anatomy_order <- c("ACC", "LFC", 'MFC', 'FP', 'OFC','SM',
                                            'INS', 'ITG', 'MTG', 'STG', 'TP', 'HPC', 'PHG', 'FG',
                                            'IPL', 'SPL', 'PMC', 'MCC',
                                            'OCC',
                                            'AMY', 'BG', 'CLT', 'antTH', 'midTH', 'pstTH')
JP_label = anatomy_order
```

### Within-subject level testing:

#### outflow pathway
```{r}
# reorganize data
ROIs = c('antTH','midTH','pstTH')
dTH <- data.frame()

for(roi in ROIs){

din_roi<- d0 %>% filter(JP_label_in != roi & JP_label_out == roi) %>% 
  mutate(elec_ID = paste(subject, record_chan), JP_label = JP_label_in, Net1 = Yeo7_in1, Net2 = Yeo7_in2, pathway = 'outflow', ROI = roi) %>% 
  select(subject, aSubID, ROI, pathway, elec_ID, JP_label,osc_speed, min_pk_time, Net1, Net2, activated, activated_default, activated_SimRule, CECS, CECS_activation, record_hot, stim_hot) 

dout_roi <- d0 %>% filter(JP_label_in == roi & JP_label_out != roi) %>% 
  mutate(elec_ID = paste(subject, stim_chan), JP_label = JP_label_out, Net1 = Yeo7_out1, Net2 = Yeo7_out2, pathway = 'inflow', ROI = roi) %>% 
  select(subject, aSubID, ROI,pathway, elec_ID, JP_label, osc_speed, min_pk_time, Net1, Net2, activated, activated_default, activated_SimRule, CECS, CECS_activation, record_hot, stim_hot)  

dTH <- rbind(dTH, din_roi, dout_roi)  
}
```



```{r}
dTH_mean <- dTH %>% filter(JP_label == 'SM') %>% 
  group_by(ROI, elec_ID) %>% 
  summarize(cecs = mean(CECS),
            sd = sd(CECS, na.rm=TRUE)) %>% 
  ungroup()

dA_mean <- dTH_mean %>% filter(ROI == 'antTH', )
dP_mean <- dTH_mean %>% filter(ROI == 'pstTH')
dmerge <- merge(dA_mean, dP_mean, by = 'elec_ID')
cecs_diff <- dmerge$cecs.y - dmerge$cecs.x

km.res <- kmeans(cecs_diff, 2, nstart=3)
df_cluster <- data.frame(cecs_diff = cecs_diff, cluster = km.res$cluster, elec_ID=dmerge$elec_ID) %>% 
  mutate(cluster_cat = case_when(
    cluster == 1 ~ '1',
    cluster == 2 ~ '2'
  ))

ggplot(df_cluster, aes(cecs_diff, fill = cluster_cat)) + geom_density(alpha = 0.2)+
  theme_classic2()

#write.csv(df_cluster, paste0(output_folder, '/R_stats/intermediate/', 'TH_PMC_cscwithclusters.csv'))

```


```{r test the cecs_diff relationship with self-hot}
dTH_PMC <- dTH %>% filter(JP_label == 'PMC')  %>% 
  mutate(hot_temp = 
               case_when(pathway=='outflow' ~ record_hot,
                         pathway=='inflow' ~ stim_hot
                           )) %>%
  mutate( hot = case_when(is.na(hot_temp) ~ 0, !is.na(hot_temp) ~ hot_temp)) %>% 
  select(-hot_temp) %>% 
  mutate(hot_cat = case_when(
    hot==1~'hot',
    hot==0~'nohot'
  ))

```

```{r}

dTH_PMC_mean <- dTH_PMC %>% 
  group_by(ROI, elec_ID) %>% 
  summarize(cecs = mean(CECS),
            sd = sd(CECS, na.rm=TRUE)) %>% 
  ungroup()


```


```{r outflow pathway, individual test}
area1 <- Area1$JP_label[Area1$nSubs>1]
path = 'outflow'
sigarea1 <- data.frame(area=c(), P=c(), Estimate=c(),tvalue=c())

for (jplabel in area1){
sbjs <- unlist(commonSbList1[which(Area1$JP_label == jplabel)])
dTH_ <- dTH %>% filter(aSubID %in% sbjs & JP_label==jplabel & ROI!='midTH' & pathway==path)
ml0 <- lmer(CECS ~ (1|subject) + (1|elec_ID), dTH_)
ml <- lmer(CECS ~ ROI+(1|subject)+ (1|elec_ID), dTH_)
fprintf('>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>')
fprintf('\nStats for %s pathway CCEP to %s between antTH and pstTH', path, jplabel)

resANOVA <- anova(ml0, ml)
resML <- summary(ml)
print(resML)
print(resANOVA)
coeff <- resML$coefficients
est<- coeff[grepl('ROI',rownames(coeff)), colnames(coeff)=='Estimate']
t <- coeff[grepl('ROI',rownames(coeff)), colnames(coeff)=='t value']
sigarea1<- rbind(sigarea1, 
                 data.frame(area=jplabel,
                            Estimate=est,
                            tvalue=t, 
                            P=resANOVA$`Pr(>Chisq)`[2]
                            ))
}
sigarea1<- sigarea1 %>% mutate(P.adj = p.adjust(P, method="fdr"))

```


#### inflow pathway

```{r inflow pathway, individual test}
area2 <- Area2$JP_label[Area2$nSubs>1]
sigarea2 <- data.frame(area=c(), P=c(), Estimate=c(),tvalue=c())
path = 'inflow'
for (jplabel in area2){ 
sbjs <- unlist(commonSbList2[which(Area2$JP_label == jplabel)])
dTH_ <- dTH %>% filter(aSubID %in% sbjs & JP_label==jplabel & ROI!='midTH' & pathway==path)
ml0 <- lmer(CECS ~ (1|subject)+ (1|elec_ID), dTH_)
ml <- lmer(CECS ~ ROI+(1|subject)+ (1|elec_ID), dTH_)
fprintf('>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>')
fprintf('\nStats for %s pathway CCEP from %s between antTH and pstTH', path, jplabel)
resANOVA <- anova(ml0, ml)
resML <- summary(ml)
print(resML)
print(resANOVA)
coeff <- resML$coefficients
est<- coeff[grepl('ROI',rownames(coeff)), colnames(coeff)=='Estimate']
t <- coeff[grepl('ROI',rownames(coeff)), colnames(coeff)=='t value']
sigarea2<- rbind(sigarea2, 
                 data.frame(area=jplabel,
                            Estimate=est,
                            tvalue=t, 
                            P=resANOVA$`Pr(>Chisq)`[2]
                            ))

}

sigarea2<- sigarea2 %>% mutate(P.adj = p.adjust(P, method="fdr"))

```


```{r, fig.width=2, fig.height=2}
sigarea<- sigarea2
area2 <- Area2$JP_label[Area2$nSubs>1]
path = 'inflow'
for (jplabel in area2){
sbjs <- unlist(commonSbList2[which(Area2$JP_label == jplabel)])
dTH_ <- dTH %>% filter(aSubID %in% sbjs & JP_label==jplabel & ROI!='midTH' & pathway==path)
dTH_mean <- dTH_ %>% 
  group_by(ROI, aSubID) %>% 
  summarize(average = mean(CECS),
            sd = sd(CECS, na.rm=TRUE)) %>% 
  ungroup()
# create plot using ggplot() and geom_boxplot() functions
ttl <- sprintf('Within-individual CECS comparison\n(inflow from %s)', jplabel)
# annotations
anot<-sprintf("P(>Chisq) = %.04f\nP(FDR-adj) = %.03f\nContrast = %.02f\nT-value = %.02f\nN(elec) = %d", 
                sigarea$P[sigarea$area==jplabel], sigarea$P.adj[sigarea$area==jplabel], 
              sigarea$Estimate[sigarea$area==jplabel], sigarea$tvalue[sigarea$area==jplabel],
              Area1$nElec[Area1$JP_label==jplabel])
xpos <- 0.5
grob <- grobTree(textGrob(anot, x=xpos,  y=0.9,
  gp=gpar(col="black", fontsize=7, fontface="italic")))

dodge = 0.3
## plot
p <- ggplot(dTH_, mapping = aes(ROI, CECS)) +
  geom_violin(color = "gray20", trim = FALSE, 
              draw_quantiles = c(0.5), linetype='dashed')+
  
  # geom_point() is used to make points at data values
  # fill and size parameters are used to customize point

  geom_pointrange(data = dTH_mean, 
             mapping = aes(x= ROI, y = average, fill = aSubID, ymin = average-sd, ymax = average+sd),
             size=0.8,shape=21, alpha = 0.7, position=position_dodge(width=dodge))+
  geom_line(data = dTH_mean, 
            mapping = aes(x = ROI, y = average, group = aSubID), 
            position=position_dodge(width=dodge), size = 0.6, color = 'grey80')+
  geom_point(aes(color=aSubID), position=position_dodge(width=dodge), size=0.5,alpha=0.6) + 
  # annot
   ylim(0, max(dTH_$CECS)*1.5)+
 annotation_custom(grob)+
  ggtitle(ttl)+
  theme_minimal() 

print(p)

# #save plot
# outDir <- paste0(figpath,'/R_stats')
# figname<-sprintf('%s/CEC %s with %s compare ant_pst_withinSbj.png', outDir, path, jplabel)
# png(figname,
#     width     = 4,
#     height    = 4.1,
#     units     = "in",
#     res       = 300,
#     pointsize = 4)
# print(p)
# dev.off()


}
```



### Test interaction effect between ROI and pathway

```{r}
# test interaction
area3 <- Area$JP_label[Area$nSubs>1]

sigarea3 <- data.frame(area=c(), P=c(), Estimate=c(),tvalue=c())

for (jplabel in area3){
sbjs <- unlist(commonSbList[which(Area$JP_label == jplabel)])
dTH_ <- dTH %>% filter(aSubID %in% sbjs & JP_label==jplabel & ROI!='midTH')
ml0 <- lmer(CECS ~ ROI+pathway+(1|subject)+ (1|elec_ID), dTH_)
ml <- lmer(CECS ~ ROI*pathway+(1|subject)+ (1|elec_ID), dTH_)
fprintf('>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>')
fprintf('\nStats for %s pathway CCEP to %s between antTH and pstTH', path, jplabel)

resANOVA <- anova(ml0, ml)
resML <- summary(ml)
print(resML)
print(resANOVA)
coeff <- resML$coefficients
est<- coeff[grepl(':',rownames(coeff)), colnames(coeff)=='Estimate']
t <- coeff[grepl(':',rownames(coeff)), colnames(coeff)=='t value']
sigarea3<- rbind(sigarea3,
                 data.frame(area=jplabel,
                            Estimate=est,
                            tvalue=t,
                            P=resANOVA$`Pr(>Chisq)`[2]
                            ))
}
sigarea3 <- sigarea3 %>% mutate(P.adj = p.adjust(P, method="fdr"))
# print(sigarea3)
# write.csv(sigarea3, paste0(output_folder, '/R_stats/sigTest_CECS_lmer_WithinSubjCmp_antVSpst_ROIpathInter.csv'))
```



